---
title: "Sanger_Psba"
author: "Magena Marzonie"
date: "11/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sangerseqR)
library(bioseq)
library(DECIPHER)
library(phangorn)
library(tidyverse)
``` 

```{r}
read_fasta_df <- function(file = ""){
    fasta <- readLines(file)
    ind <- grep(">", fasta)
    s <- data.frame(ind = ind, from = ind + 1, to = c((ind - 
        1)[-1], length(fasta)))
    seqs <- rep(NA, length(ind))
    for (i in 1:length(ind)) {
        seqs[i] <- paste(fasta[s$from[i]:s$to[i]], collapse = "")
    }
    tib <- tibble(label = gsub(">", "", fasta[ind]), 
        sequence = seqs)
    return(tib)
}

write_fasta_df <- function (data, filename) 
{
    fastaLines = c()
    for (rowNum in 1:nrow(data)) {
        fastaLines = c(fastaLines, as.character(paste(">", 
            data[rowNum, "label"], sep = "")))
        fastaLines = c(fastaLines, as.character(data[rowNum, 
            "sequence"]))
    }
    fileConn <- file(filename)
    writeLines(fastaLines, fileConn)
    close(fileConn)
}

dna_to_DNAbin <- function (dna){
  DNAbin <- as_DNAbin(dna)
  names(DNAbin) <- names(dna)
  return(DNAbin)
}
dna_to_DNAStringset <- function(x) 
{
    bioseq:::check_dna(x)
    DNAstr <- DNAStringSet(paste(x))
    names(DNAstr) <- names(x)
    return(DNAstr)
}

DNAStringSet_to_dna <- function(x){
    x_dna <- as_dna(paste(x))
    names(x_dna) <- names(x)
    res <- tibble(label = names(x), sequence = x_dna)
    return(res)
}

gg_chromatogram <- function (file = "", row_number = 4, ratio = 0.33) 
{
    calls <- makeBaseCalls(readsangerseq(file), ratio = 0.33)
    peak_length <- max(calls@peakPosMatrix, na.rm = TRUE)
    seq <- primarySeq(calls)
    seq_length <- str_length(seq)
    factor <- peak_length/seq_length
    trace <- calls@traceMatrix %>% as.data.frame() %>% select(A = V1, 
        C = V2, G = V3, T = V4) %>% mutate(pos = row_number()) %>% 
        tidyr::pivot_longer(A:T, names_to = "Nucleotide", 
            values_to = "Intensity") %>% 
      mutate(facet = rep(x = LETTERS[1:row_number], 
        each = ceiling(n()/row_number))[1:n()])

  long_seq <- str_split(seq, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(pos = row_number(),
         L1 = as.character(L1),
         y = max(trace$Intensity),
         Nucleotide = value) %>%
    mutate(facet = rep(x = LETTERS[1:row_number], 
        each = ceiling(n()/row_number))[1:n()])
    
    theme_set(theme_bw())
    palette <- c(A = "#6bb04a", G = "#4e76dd", C = "#c89232", T = "#d14d4b")
    
    ggplot() + 
      geom_text(aes(x = pos, y = y, colour = Nucleotide, label = Nucleotide), na.rm = TRUE, long_seq) + 
      geom_path(aes(x = pos/factor, y = Intensity, group = Nucleotide, colour = Nucleotide), trace) + 
      facet_wrap(~facet, ncol = 1, scales = "free_x") + 
      theme(strip.background = element_blank(), strip.text.x = element_blank()) + 
      scale_color_manual(values = palette) + ylab("Intensity") + 
      scale_x_continuous(breaks = scales::breaks_width(50)) +
      xlab("Position")
}
```

# Viewing chromatograms

Use this plot to help decide where the trimming start and stop positions should be

```{r}
gg_chromatogram(file = "/Users/magenamarzonie/Library/CloudStorage/OneDrive-JamesCookUniversity/PhD Chapter 2/Psba Sanger 2022/Raw_Psba_Files/psba1_A01_74_Forw.ab1", row_number = 5, ratio = 0.33)

gg_chromatogram(file = "/Users/magenamarzonie/Library/CloudStorage/OneDrive-JamesCookUniversity/PhD Chapter 2/Psba Sanger 2022/Raw_Psba_Files/psba1_A01_78_Rev.ab1", row_number = 5, ratio = 0.33)
```

# importing all the forward .fa or .txt files (same thing)
```{r}
f_all <- sort(list.files(path = "/Users/magenamarzonie/Library/CloudStorage/OneDrive-JamesCookUniversity/PhD Chapter 2/Psba Sanger 2022/Raw_Psba_Files/", pattern = "_74_Forw.txt", full.names = TRUE))
r_all <- sort(list.files(path = "/Users/magenamarzonie/Library/CloudStorage/OneDrive-JamesCookUniversity/PhD Chapter 2/Psba Sanger 2022/Raw_Psba_Files/", pattern = "_78_Rev.txt", full.names = TRUE))
```

```{r}
# If this map & reduce code chunk produces the error: ' incomplete final line found on ..', run this code first to add \n to end of each txt file. WARNING - this modifies the original final! Create backups first.

for(i in 1:length(f_all)){
   cat("\n", file = f_all[i], append = TRUE)
   cat("\n", file = r_all[i], append = TRUE)
 }

f_all_fasta <- f_all %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe

r_all_fasta <- r_all %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe
```

# apply the trimming parameters to start and end of the sequences (example, remove 10 bases from start and end)

```{r}
f_all_fasta_sub <- f_all_fasta %>%
  mutate(sequence = str_sub(sequence, start = 60, end = 830)) # start = trim + 1, end = trim + 1

r_all_fasta_sub <- r_all_fasta %>%
  mutate(sequence = str_sub(sequence, start = 50, end = 700)) # start = trim + 1, end = trim + 1
```

# convert to format for DECIPHER and reverse complement the reverse sequence

```{r}
f_dss <- f_all_fasta_sub %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset()

r_dss <- r_all_fasta_sub %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset() %>%
  reverseComplement()

pair_list <- list()
for(i in 1:length(f_dss)){
  ss <- DNAStringSet(c(f_dss[i], r_dss[i]))
  pair_list[[i]] <- ss
}
```

# Align the sample pairs

```{r}
alignment_list <- list()
for(i in 1:length(pair_list)){
  alignment <- AlignSeqs(pair_list[[i]])
  alignment_list[[i]] <- alignment
}
```

# View the alignments

```{r}
palette <- c("A" = "#6bb04a", "G" = "#4e76dd", "C" = "#c89232", "T" = "#d14d4b", "N" = "black", "-" = "white", "K" = "purple", "variable_site" = "grey60", "invariable_site" = "grey90")

aligned_df <- data.frame()
for(i in 1:length(alignment_list)){
  a_pair <- alignment_list[[i]] %>% writeXStringSet("temp_file.fasta")
  a_df <- read_fasta_df("temp_file.fasta")
  aligned_df <- rbind(aligned_df, a_df)
}

aligned_plotting <- aligned_df %>%
  mutate(sample_id = str_sub(label, 23, 25)) # create a metadata column that identifies the sample

# Create profile key
key <- aligned_plotting %>%
  tibble::rownames_to_column(var = "id")

# Create long dataframe for ggplot
long_sequences <- str_split(aligned_plotting$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

# Plot alignment
ggplot(long_sequences, aes(y = label, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      facet_wrap(~ sample_id, ncol = 1, scales = "free_y") +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank()) +
      scale_x_continuous(expand = c(0, 0)) +
    xlab("Position")

```

# Create the consensus sequences

```{r}
consensus_df <- list()
for(i in 1:length(alignment_list)){
  a_con <- alignment_list[[i]] %>% ConsensusSequence()
  names(a_con) <- names(alignment_list[[i]])[1]
  consensus_df[[i]] <- a_con
}
```

# Align the consensus sequences

```{r}
aligned_consensus <- AlignSeqs(do.call(c, consensus_df))
writeXStringSet(aligned_consensus, "consensus_sequence_alignment.fasta")
```

# Visualise the alignment

```{r}
final_alignment <- read_fasta_df("consensus_sequence_alignment.fasta") %>%
  mutate(sample_id = str_sub(label, 7, 10)) # create a metadata column that identifies the sample

# Create profile key
final_key <- final_alignment %>%
  tibble::rownames_to_column(var = "id") %>%
  mutate(sample_id = str_sub(label, 23, 25))

# Create long dataframe for ggplot
final_long_sequences <- str_split(final_alignment$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., final_key, by = c("L1" = "id")) %>%
  ungroup()

matrix_block <- final_long_sequences %>%
  group_by(x) %>%
  mutate(variant = case_when(length(unique(value)) > 1 ~ "variable_site",
                             length(unique(value)) == 1 ~ "invariable_site")) %>%
  sample_n(1) %>%
  mutate(label = "variable_sites")

n1 <- length(unique(final_long_sequences$x))
n2 <- length(unique(final_long_sequences$sample_id))

# Plot alignment
ggplot(final_long_sequences, aes(y = sample_id, x = x)) +
      geom_tile(aes(fill = value), height = 1) +
      geom_tile(aes(x = x, y = label, fill = variant), height = 1, matrix_block) +
      scale_fill_manual(values = palette, name = "base") +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank()) +
      scale_x_continuous(expand = c(0, 0)) +
   geom_line(data = data.frame(x = c(0, n1) + 0.5, y = rep(2:n2, each = 2) - 0.5),
            aes(x = x, y = y, group = y)) + # Horizonal lines
  xlab("Position")
```


